Accurate binding of calcium to phospholipid bilayers by effective inclusion of electronic polarization

### 2017-08-08 Manuscript/manuscript.tex
Manuscript update after group discussion. Concentration corrected to actual values, not nominal. Ord.Pars include exp. values for tails. Updated Ca-density, included water profile.

% Here I write pseudo-article statements that will make the main argument.
% Beautiful polished sentences will be formed only afte we agree on these basic things
\section{Introduction}
Cation interactions with cellular membranes play a key role in several biological processes,
like in signal propagation in neurons and vesicle fusion.
\todo{JOE: list more introductory examples, where it has interesting effects.}
Zwitterionic phosphocholine (PC) lipid bilayers have been widely used as model systems
to understand molecular level details of specific cation interactions with
cellular membranes by using experimental \cite{cevc90,tocanne90,binder02,pabst07,uhrikova08}
and theoretical methods \cite{bockmann03,bockmann04,Berkowitz12,??} \todo{We need to select references for these}.
While relative binding affinity of different ions is agreed to follow Hoffmeister
series, the molecular details of binding and binding energetics are
not fully understood \cite{cevc90,tocanne90,seelig90,binder02,catte16,??}.
%General conclusion from experimental
%studies has been that multivalent ions and lithium have weak specific binding in
%phopholipid bilayers, while other monovalent ions do not essentially
%bind 
%The presence of anoinic lipids, like PS or PG,
%increase the concentration close to the bilayer and thus the amount of bound
%ions, but do not affect the specific binding constant \cite{seelig90}.
Non-invansive spectroscopic methods, like nucelar magnetic resosnance (NMR), scattering
and infrared scpectroscopy, give accurate information about ion binding
in lipid bilayers\cite{hauser76,hauser78,herbette84,akutsu81,altenbach84,binder02}. 
Molecular level interpretation of the results, however, requires assumptions
about binding model and is often not fully conclusive \cite{??}.
On the other hand, the accuracy of the state of the art atomistic resolution lipid and ion models
have turned out insufficient for the detailed interpretation of cation binding details \cite{catte16}.
The overestimated specific cation binding observed in current lipid models \cite{catte16}
may lead to articifially positively charged memranes and significant artefacts in
MD simualtions (e.g.~\ce{Ca^2+} induced charge inversion of bacterial membranes~\cite{Luan2016}).

@@

In this work we show that the accuracy of ion binding, especially calcium,
in zwitterionic PC lipid bilayer can be significantly improved 
for classical MD simulation models of zwitterionic lipids
by including electronic polarizability
in the polar region of lipid molecules
using the electronic continuum correction (ECC)~\cite{leontyev11}. 
This approach has been previously shown to improve performace of
ion models against neutron scattering data in concentrated solutions \cite{Jungwirth2015,kohagen14,kohagen16}
and ab-initio simulations~\cite{Jungwirth2017}. 
The realistic structure of ions in a bulk solvent was not, however, sufficient to correct binding in lipid
bilayers~\cite{catte16}. 
To fix it, we propose
a classical fixed-charge MD simulation model of 
1-Palmitoyl-2-oleoylphosphatidylcholine (POPC) lipid 
that accounts for electronic polarizability by using ECC. 
We will validate such a model by comparing its lipid bilayer structue
to x-ray scattering data and its head group structure to NMR experiments. 
We will then quantify the accuracy of head group order parameter response to bound charge
and compare it to the experimental NMR data of sodium and calcium binding in PC lipid bilayers. 
In addition, we will quantify relative binding affinities of different moieties in POPC to such cations. 
This will be discussed in terms of binding modes, stoichiometry and binding isotherms and their models used in experiments.
The proposed POPC model, ECC-lipids, and the approach to obtain it
is, hence, highly useful for future MD simulations with physiological salt conditions, 
and for effects where membrane electrostatics play a major role.

@@

While the scaling factor of $f_q = 0.75$ for ions in water is well justified and shown
to improve model performance against scattering data \cite{kohagen14,kohagen16, Pluharova2014},
it is not clear if the same factor should be used for lipid head groups. 
Unlike the total charge of cations and anions, atomic partial charges used in
classical MD simulations are not experimental observables and 
can in principle adopt arbitrary values. 
Here we estimate the correct scaling factor for Lipid14 model by using
so called ``implicitly polarized charges'' (IPolQ) \cite{ipolq2013},
which are the average of partial charges given by 
restrained electrostatic potential fitting (RESP~\cite{RESP_paper}) 
in vacuum and in solvent.
Using the partial atomic charges obtained from RESP fitting in vacuum and implicit solvent reported in \cite{maciejewski14} and
considering that the partial charges in original Lipid14 parametrization were derived using RESP~\cite{RESP_paper} in vacuum,
we conclude that the charges in Lipid14 model shall be scaled roughly by the factor $f_q=0.8$. 
%By taking the charges of oxygen atoms in vacuum and implicit water solvent for POPC from \cite{maciejewski14}, 
%we represent the IPolQ charges in the electronic continuum correction of the Lipid14 model
%by increasing the scaling factor $f_q$ to $0.8$. 
%-----\footnote----------
  %{Depending on which QM method you arrive at values from 0.76 to 0.83, averages across atom
  %types being around 0.78--0.80. Even though the methods are almost identical, authors of Lipid 14 find lower
  %partial charges in vacuum than here -- so I prefer the higher value. As the choice of charges is arbitrary anyway,
  %I use 0.80 as an approximate round value, which accounts for the already lower charges in Lipid14. 
  %The use of 0.78---.79 might be more appropriate, though.}

Here we apply the ECC correction with the proposed scaling factor only to the head group,
glycerol backbone and carbonyl regions, which are the most polar parts in lipids and
are expected to have the largest contribution to the cation binding.
The hydrocarbon chain parameters are not modified, because they are
already highly optimized in most lipids models including Lipid14 and
give generally a good description for hydrophobic part of lipid bilayers
in various conditions \cite{ollila16}.
In contrast, all available lipid models require
some improvements in glycerol backbone and head group regions \cite{botan15}. 
%To minimize the detuning of the highly optimized hydrocarbon
%parameters, 
%Because the biologically relevant cations like \ce{Na^+} and \ce{Ca^{2+}}
%are found to interact only with the phospholipid head group,
%and the accuracy of alkyl sidechains in Lipid14 model are satisfactory, 
%we exclude alkyl tails from our modifications. 
The scaling of partial charges in the polar region was found to decrease
area per molecule to 60~\AA$^2$, which is significantly smaller than the
experimental value 64.3~\AA$^2$ (\cite{}\todo{missing REF for APL experiment}) and the original Lipid14 values $(65.6 \pm 0.5)$~\AA$^2$~\cite{dickson14}..  
The decrease of area was found to arise from a lower hydration of the lipid head group region,
%\todo{SAMULI: Did you analyze the effect of this to hydration or did we only speculate?
%JOE: I found a systematic decrease of water density in the heagroup region even if I kept APL constant.}
which can be explained by the decreased solvation free energy due to the lower polarity
of molecules with scaled charges. 
The solvation free energy and hydration is sensitive to Van der Waals interaction parameters,
hence, they can be increased by reducing the radius of the selected atoms by changing the 
the $\sigma$ parameters in Lennard-Jones potential. 
%This allows water molecules to approach closer to lipid atoms and
%have stonger electrostatic interactions with them.
%\todo{SAMULI: This effect may have an official name. In that case we should mention it.}
Here we reduce $\sigma$ with a scaling factor of $f_\sigma = 0.89$ for the same atoms for which
charges were scaled. This has lead to an increase of the area per molecule back
close to the experimental value (see Table~\ref{tab:apls}). 
%\todo{Is there some justification for the below sentence?} 
%In line with the electronic continnum correction,
%the scaling factor for sigma parameters, $f_\sigma$,
%has to lay between $f_q$ and 1.

@@

%\todo{ongoing,Actual concentration of cations in simulation has yet to be estimated. 
%If it varies too much from the nominal concentration, I may need to tweak the scaling 
%factors, $f_q$ or only $f_\sigma$, to accomodate it. However, it is very unlikely, response 
%to the surfactant \ce{DHMDMAB} is OK. Big patches with loads of solvent are running at the 
%moment to guide me on the possible finite-size errors and this conc-error.
The used experimental data report order parameters as a function of
equilibrium cation concentration in the bulk solvent \cite{akutsu81,altenbach84}.
Such a condition is reached in simulations by adjusting
the simulation box size to dimensions large enough
that ion conceration reaches a clear plateau in the bulk solvent.
The concentrations in the units of $\mathrm{mol/l}$ were then 
determined as
\begin{equation}
  C_{eq}=\frac{C_{plateau}}{0.602},
\end{equation}
where plateau concentration is the number density in the units of $\mathrm{nm}^{-3}$.
\todo{SAMULI: Once we have to final results, we can probably say that
  the repeat distance is not far from the experimentally measured distance \cite{pabst07,uhrikova08}
  }

\subsection{Comparison of lipid bilayer structure to experimental data}

@@

\begin{figure}[tbp]
  \centering
  \includegraphics[height=8.5cm,angle=-90]{../Fig/form-f_exp-l14-eccl17.eps}
  \includegraphics[width=9.0cm]{../Fig/ipython_nb/Order-parameters_exp-L14-ECCL17_q80_sig89.png}
  \caption{\label{simVSexpNOions}
    X-ray scattering form factors from experiments~\cite{Kucerka2011} and simulations using Lipid14 \cite{dickson14} and ECC-lipids models. 
    Order parameters of head group, glycerol backbone and sn-1 and sn-2 tails  from simulations with Lipid14 \cite{dickson14} and ECC-lipids models
    compared with experimental order parameters from \cite{ferreira13}.}
  \todo{Add acyl chain order parameters, POPC chemical structure} \\
  \todo{We should run a simulation for the system withtout ions with the
    same (or close) temperature as the experimental data. The NMR data is at 300K
    so that would be good. I think that this would be also close enough for the scattering data at 303K.
  }
\end{figure}

@@

In order to validate the newly developed model, ECC-lipids, 
we compared our simulation results without ions to
NMR order parameters measurements and x-ray scattering form factors 
(Fig.~\ref{simVSexpNOions} and Table~\ref{tab:apls}). 
The tail order parameters being already highly optimized in the original Lipid14 model~\cite{dickson14}
are found to match the experimental values nicely.
The head group and glycerol backbone order parameters accuracy is 
comparable to the state of art lipid models available in literature \cite{ollila16},
The agreement between the x-ray scattering form factors and areas per molecule from simulations and experiments
confirm that the membrane structural properties are well captured. 
%The optimal value for $f_\sigma$ was found by representing the overall membrane structure well
%by matching scattering form factors to experimental data~\cite{Petrache06, Kucerka08, Pabst10}.

@@

\begin{figure}[tbp]
  \centering
  \includegraphics[width=9.0cm,angle=0]{../Fig/CAdensities.eps}
  \caption{\label{fig:cacl-dens}
    Density profiles of \ce{Ca^{2+}} and \ce{Cl^-} for Lipid14 model with Dang ions
    and with ECC-ions and ECC-lipids model with ECC-ions. }
  \todo{Change the figure so that it contains a membrane background} \\
  \todo{What are the concentrations (calculated as in NMRlipids II) in this figure?} \\
  \todo{These should be also compared to the results with Åqvist ions, which are shown
    in NMRlipids II. NMRlipids II have results with 350mM and 1000mM. If we have the distributions with
    the same concentrations Åqvist results could be added also in this figure. If not, the
  comparison has to be made some other way.}
\end{figure}

After achieving accurate head group order parameter response with cationic surfactant, 
we advanced to the response to solutions with varying concentrations of salts, namely \ce{NaCl} and \ce{CaCl_2}. 
Headgroup order parameter response to increasing CaCl$_2$ concentration
from experiments, original Lipid14 model and ECC-lipids model are shown
in Fig. \ref{OrderParameterCHANGESnewMODELS}. The order parameter response
is significantly overestimated in original Lipid14 model, while results from
ECC-lipids model are in a good agreement with the experiments.
This is a significant improvement over current models,
which in general overstimate the head group order parameter response to CaCl$_2$
concentration \cite{catte16}. The good agreement with experiments indicate
that the binding details of \ce{Ca^2+} are realistic in the ECC-lipids model
and, hence, it can be used to study lipid-ion interaction in detail.

Ion density profiles from simulations with different models are compared in Fig.~\ref{fig:cacl-dens}.
The density profiles from simulations with original Lipid14 \cite{dickson14} and Dang ions \cite{smith94,chang1999,dang2006} 
show a pronounced peak in the position of the phosphate moieties of POPC. 
The use of a ECC-ion model~\cite{kohagen14,kohagen16} 
or the standard Amber ion model by \AA{quist} 
along with Lipid14 does not significantly change it 
(Headgroup order parameter responses for these models in SI).
\todo{Add OP-response of Lipid14+ECC-ions into the OP-changes plot in SI}
The new ECC-lipids model with ECC-ions exhibits on the other hand smaller density in this region 
suggesting overall weaker binding of cations (Fig.~\ref{OrderParameterCHANGESnewMODELS}). 
Together with the accurate head group order parameter response (Fig.~\ref{OrderParameterCHANGESnewMODELS}),
we can conclude that the present model, ECC-lipids, exhibits accurate ion binding affinity to zwitterionic phospholipid membranes
suggesting that the models that overestimate head group order parameter response~\cite{catte16} also overestimate the amount of bound cations.

\begin{figure}[]
  \centering
  \includegraphics[height=9.0cm,angle=-90]{../Fig/bound-CAs_conc-eccl17.eps}
  \caption{\label{fig:cacl-bind}
    Binding isotherm assuming stoichiometry of 2~POPC:1~\ce{Ca^{2+}} as used in \cite{altenbach84} fits our simulation data well .}
\end{figure}

Good agreement of ECC-lipids model with experiments encourages us to analyse the
binding details from MD simulations. Direct analysis of contacts between ions and
lipids from simulations suggest that the most abundant POPC:\ce{Ca^2+} complex 
have stoichiometry of 2~POPC:1~\ce{Ca^{2+}}.
As shown in Fig.~\ref{fig:cacl-bind} this is in agreement
with the ternary complex model suggested based on head group order parameter
experiments~\cite{altenbach84}.
In addition to the ternary complexes, there also is a non-negligable probability
of one \ce{Ca^{2+}} cross-bridging three POPC molecules.
Technical details of the analysis are in the SI.

@@

The electronic continuum correction is applied here on Lipid14 POPC model \cite{dickson14},
but we expect that the correction can be generalized also for other lipids
and force fields.
The parameters can be used with existing standard nucleic acid and protein force fields, e.g. AMBER-FB15~\cite{Wang2017}.

>>>>

% Here I write pseudo-article statements that will make the main argument.
% Beautiful polished sentences will be formed only afte we agree on these basic things
\section{Introduction}
Cation interactions with cellular membranes play a key role in several biological processes,
like in signal propagation in neurons and vesicle fusion.
\todo{JOE: list more introductory examples, where it has interesting effects.}
Zwitterionic phosphocholine (PC) lipid bilayers have been widely used as model systems
to understand molecular level details of specific cation interactions with
cellular membranes by using experimental \cite{cevc90,tocanne90,binder02,pabst07,uhrikova08}
and theoretical methods \cite{bockmann03,bockmann04,Berkowitz12,??} 
\todo{PAVEL: introduce previous theoretical work that discusses cation binding to POPC w.r.t. its specific moiteties, e.g. Lukas' paper}.
While relative binding affinity of different ions is agreed to follow Hoffmeister
series, the molecular details of binding and binding energetics are
not fully understood \cite{cevc90,tocanne90,seelig90,binder02,catte16,??}.
%General conclusion from experimental
%studies has been that multivalent ions and lithium have weak specific binding in
%phopholipid bilayers, while other monovalent ions do not essentially
%bind 
%The presence of anoinic lipids, like PS or PG,
%increase the concentration close to the bilayer and thus the amount of bound
%ions, but do not affect the specific binding constant \cite{seelig90}.
Non-invansive spectroscopic methods, like nucelar magnetic resosnance (NMR), scattering
and infrared scpectroscopy, give accurate information about ion binding
in lipid bilayers\cite{hauser76,hauser78,herbette84,akutsu81,altenbach84,binder02}. 
Molecular level interpretation of the results, however, requires assumptions
about binding models and is often not fully conclusive \cite{??}.
On the other hand, the accuracy of the state of the art atomistic resolution lipid and ion models
have turned out insufficient for the detailed interpretation of cation binding details \cite{catte16}.
The overestimated specific cation binding observed in current lipid models \cite{catte16}
may lead to articifially positively charged memranes and significant artefacts in
MD simualtions (e.g.~divalent cation-induced charge inversion of bacterial membranes~\cite{Luan2016} \todo{JOE: this may be a too daring statement.}).

@@

In this work we show that the accuracy of ion binding, especially calcium,
in zwitterionic PC lipid bilayer can be significantly improved 
for classical MD simulation models of zwitterionic lipids
by including electronic polarizability
in the polar region of lipid molecules
using the electronic continuum correction (ECC)~\cite{leontyev11}. 
This approach has been previously shown to improve performace of
ion models against neutron scattering data in concentrated solutions 
and ab-initio simulations~\cite{Jungwirth2017,Jungwirth2015,kohagen14,kohagen16}. 
The realistic structure of ions in a bulk solvent was not, however, sufficient to correct binding in lipid
bilayers~\cite{catte16}. 
To fix it, we propose
a classical fixed-charge MD simulation model of 
1-Palmitoyl-2-oleoylphosphatidylcholine (POPC) lipid 
that accounts for electronic polarizability by using ECC. 
We will validate such a model by comparing its lipid bilayer structue
to x-ray scattering data and its head group structure to NMR experiments. 
We will then quantify the accuracy of the head group order parameter response to the bound charge
and compare it to the experimental NMR data of sodium and calcium binding in PC lipid bilayers. 
In addition, we will quantify relative binding affinities of different moieties in POPC to such cations. 
This will be discussed in terms of binding modes, stoichiometry and binding isotherms and their models used in experiments.
The proposed POPC model, ECC-lipids, and the approach to obtain it
is, hence, highly useful for future MD simulations with physiological salt conditions, 
and for effects where membrane electrostatics play a major role.

@@

While the scaling factor of $f_q = 0.75$ for ions in water is well justified and shown
to improve model performance against scattering data \cite{kohagen14,kohagen16, Pluharova2014},
it is not clear if the same factor should be used for molecules, i.e. lipid head groups in our case. 
Unlike the total charge of cations and anions, atomic partial charges used in
classical MD simulations are not experimental observables and 
can in principle adopt arbitrary values. 
In cases where the partial charges have some physical meaning 
(e.g. approximate electrostatic potential as in RESP~\cite{RESP_paper}),
we can merely estimate the scaling factor for molecules, $f_q$, 
to adopt a value between $\epsilon _{el} ^{-1/2}$ and $1$.
Searching in this parameter space, we empirically discovered that
the scaling factor $f_q = 0.8$ is optimal for applying ECC on Lipid14 model. 
Such a finding is in line with the estimate considering the
so called ``implicitly polarized charges'' (IPolQ) \cite{ipolq2013}.
IPolQ charges are obtained as the average of partial charges given by 
restrained electrostatic potential fitting (RESP~\cite{RESP_paper}) 
in vacuum and in a solvent.
Using the partial atomic charges obtained from RESP fitting in vacuum and implicit solvent reported in \cite{maciejewski14} and
considering that the partial charges in original Lipid14 parametrization were derived using RESP~\cite{RESP_paper} in vacuum,
we find out that the charges in Lipid14 model shall be scaled roughly by the factor $f_q=0.8$ 
-- agreeing with our empirical finding. 
%By taking the charges of oxygen atoms in vacuum and implicit water solvent for POPC from \cite{maciejewski14}, 
%we represent the IPolQ charges in the electronic continuum correction of the Lipid14 model
%by increasing the scaling factor $f_q$ to $0.8$. 
%-----\footnote----------
  %{Depending on which QM method you arrive at values from 0.76 to 0.83, averages across atom
  %types being around 0.78--0.80. Even though the methods are almost identical, authors of Lipid 14 find lower
  %partial charges in vacuum than here -- so I prefer the higher value. As the choice of charges is arbitrary anyway,
  %I use 0.80 as an approximate round value, which accounts for the already lower charges in Lipid14. 
  %The use of 0.78---.79 might be more appropriate, though.}

Here we apply the ECC correction with the proposed scaling factor only to the head group,
glycerol backbone and carbonyl regions, which are the most polar parts in lipids and
are expected to have the largest contribution to the cation binding.
The hydrocarbon chain parameters are not modified, because they are
already highly optimized in most lipids models including Lipid14 and
give generally a good description for hydrophobic part of lipid bilayers
in various conditions \cite{ollila16}.
In contrast, all available lipid models require
some improvements in glycerol backbone and head group regions \cite{botan15}. 
%To minimize the detuning of the highly optimized hydrocarbon
%parameters, 
%Because the biologically relevant cations like \ce{Na^+} and \ce{Ca^{2+}}
%are found to interact only with the phospholipid head group,
%and the accuracy of alkyl sidechains in Lipid14 model are satisfactory, 
%we exclude alkyl tails from our modifications.

The scaling of partial charges in the polar region was found to 
significantly decrease cation binding and
the area per molecule to $\approx$60~\AA$^2$, which is significantly smaller than the
experimental value 64.3~\AA$^2$ (\cite{}\todo{missing REF for APL experiment}) and the original Lipid14 values $(65.6 \pm 0.5)$~\AA$^2$~\cite{dickson14}.
The decrease of area was found to arise from a lower hydration of the lipid head group region,
%\todo{SAMULI: Did you analyze the effect of this to hydration or did we only speculate?
%JOE: I found a systematic decrease of water density in the heagroup region even if I kept APL constant.}
which can be explained by the decreased solvation free energy due to the lower polarity
of molecules with scaled charges. 
The solvation free energy and hydration is sensitive to Van der Waals interaction parameters,
hence, they can be increased by reducing the radius of the selected atoms by changing the 
the $\sigma$ parameters in Lennard-Jones potential. 
%This allows water molecules to approach closer to lipid atoms and
%have stonger electrostatic interactions with them.
%\todo{SAMULI: This effect may have an official name. In that case we should mention it.}
Here we reduce $\sigma$ with a scaling factor of $f_\sigma = 0.89$ for the same atoms for which
charges were scaled. This has lead to an increase of the area per molecule back
close to the experimental value (see Table~\ref{tab:apls}). 
%\todo{Is there some justification for the below sentence?} 
%In line with the electronic continnum correction,
%the scaling factor for sigma parameters, $f_\sigma$,
%has to lay between $f_q$ and 1.

@@

%\todo{ongoing,Actual concentration of cations in simulation has yet to be estimated. 
%If it varies too much from the nominal concentration, I may need to tweak the scaling 
%factors, $f_q$ or only $f_\sigma$, to accomodate it. However, it is very unlikely, response 
%to the surfactant \ce{DHMDMAB} is OK. Big patches with loads of solvent are running at the 
%moment to guide me on the possible finite-size errors and this conc-error.
The used experimental data report order parameters as a function of
equilibrium cation concentration in the bulk solvent \cite{akutsu81,altenbach84}.
Such a condition is reached in simulations by adjusting
the simulation box size to dimensions large enough
that ion conceration reaches a clear plateau in the bulk solvent.
The concentrations in the units of $\mathrm{mol/l}$ were then 
determined as
\begin{equation}
\label{eq:conc}
  C_{eq}=\frac{C_{plateau}}{0.602},
\end{equation}
where plateau concentration is the number density in the units of $\mathrm{nm}^{-3}$.
\todo{SAMULI: Once we have to final results, we can probably say that
  the repeat distance is not far from the experimentally measured distance \cite{pabst07,uhrikova08}
  }

\subsection{Comparison of lipid bilayer structure to experiments}

@@

\begin{figure}[tbp]
  \centering
  \includegraphics[height=8.5cm,angle=-90]{../Fig/form-f_exp-l14-eccl17.eps}
  \includegraphics[width=7.0cm]{../Fig/POPCstructure.eps}
  \includegraphics[width=9.0cm]{../Fig/ipython_nb/Order-parameters_exp-L14-ECCL17_q80_sig89.png}
  \caption{\label{simVSexpNOions}
    X-ray scattering form factors from experiments~\cite{Kucerka2011} and simulations using Lipid14 \cite{dickson14} and ECC-lipids models. 
    Order parameters of head group, glycerol backbone and sn-1 and sn-2 tails  from simulations with Lipid14 \cite{dickson14} and ECC-lipids models
    compared with experimental order parameters from \cite{ferreira13}.}
  \todo{We should run a simulation for the system withtout ions with the
    same (or close) temperature as the experimental data. The NMR data is at 300K
    so that would be good. I think that this would be also close enough for the scattering data at 303K.
  }
\end{figure}

@@

In order to validate the newly developed model, ECC-lipids, 
we compared our simulation results without any ions to
NMR order parameters measurements and x-ray scattering form factors 
(Fig.~\ref{simVSexpNOions} and Table~\ref{tab:apls}). 
The tail order parameters being already highly optimized in the original Lipid14 model~\cite{dickson14}
are found to match the experimental values nicely.
The head group and glycerol backbone order parameters accuracy is 
comparable to the state of art lipid models available in literature \cite{ollila16}.
The agreement between the x-ray scattering form factors and the areas per molecule from simulations and experiments
confirm that the membrane structural properties are well captured. 
A structural comparison of ECC-lipids with Lipid14 can be found 
in SI along with results with other water models. 
%The optimal value for $f_\sigma$ was found by representing the overall membrane structure well
%by matching scattering form factors to experimental data~\cite{Petrache06, Kucerka08, Pabst10}.

@@

\begin{figure}[tbp]
  \centering
  \includegraphics[width=9.0cm,angle=0]{../Fig/CAdensities.eps}
  \caption{\label{fig:cacl-dens}
    Profiles of normalized number density of \ce{Ca^{2+}} and \ce{Cl^-} for Lipid14 model with Dang ions
    and with ECC-ions and ECC-lipids model with ECC-ions. 
    Normalization factor for ions is 1 for monovalent ions (i.e.~\ce{Na^+} and~\ce{Cl^-}), 2 for divalent ions (i.e.~\ce{Ca^2+}), and $1/100$ for water. 
    }
  \todo{PAVEL: draw phosphate position with its variance, add water density (scaled) and include the number of$\Gamma$-surface access.} \\
  \todo{JOE: Change the figure so that it contains a membrane background} \\
  \todo{What are the concentrations (calculated as in NMRlipids II) in this figure?} \\
  \todo{These should be also compared to the results with Åqvist ions, which are shown
    in NMRlipids II. NMRlipids II have results with 350mM and 1000mM. If we have the distributions with
    the same concentrations Åqvist results could be added also in this figure. If not, the
  comparison has to be made some other way.}
\end{figure}

After achieving accurate head group order parameter response with a cationic surfactant, 
we advanced to quantifying the response to solutions with varying concentrations of salts, namely \ce{NaCl} and \ce{CaCl_2}. 
Headgroup order parameter response to increasing CaCl$_2$ concentration
(as defined in Equation~\ref{eq:conc})
from experiments, original Lipid14 model and ECC-lipids model are shown
in Fig. \ref{OrderParameterCHANGESnewMODELS}. The order parameter response
is significantly overestimated in original Lipid14 model, while results from
ECC-lipids model are in a good agreement with the experiments.
This is a significant improvement over current models,
which in general overstimate the head group order parameter response to CaCl$_2$
concentration \cite{catte16}. The good agreement with experiments indicate
that the binding details of \ce{Ca^2+} are realistic in the ECC-lipids model
and, hence, it can be used to study lipid-ion interaction in detail.

Ion density profiles from simulations with different models are compared in Fig.~\ref{fig:cacl-dens}.
The density profiles from simulations with original Lipid14 \cite{dickson14} and Dang ions \cite{smith94,chang1999,dang2006} 
show a pronounced peak in the position of the phosphate moieties of POPC. 
The use of a ECC-ion model~\cite{Jungwirth2017,Jungwirth2015,kohagen14,kohagen16}
or the standard Amber ion model by \AA{qvist}~\cite{aqvist90}
along with Lipid14 does not significantly change it 
(Headgroup order parameter responses for these models in SI).
\todo{Add OP-response of Lipid14+ECC-ions into the OP-changes plot in SI}
The new ECC-lipids model with ECC-ions~\cite{Jungwirth2017,Jungwirth2015,kohagen14,kohagen16} exhibits on the other hand smaller density in this region 
suggesting overall weaker binding of cations (Fig.~\ref{OrderParameterCHANGESnewMODELS}). 
Together with the accurate head group order parameter response (Fig.~\ref{OrderParameterCHANGESnewMODELS}),
we can conclude that the present model, ECC-lipids, exhibits accurate ion binding affinity to zwitterionic phospholipid membranes
suggesting that the models that overestimate head group order parameter response~\cite{catte16} also overestimate the amount of bound cations.

\begin{figure}[]
  \centering
  \includegraphics[height=9.0cm,angle=-90]{../Fig/bound-CAs_conc-eccl17.eps}
  \caption{\label{fig:cacl-bind}
    Binding isotherm assuming stoichiometry of 2~POPC:1~\ce{Ca^{2+}} as used in \cite{altenbach84} fits our simulation data well .
    }
\end{figure}

The good agreement of ECC-lipids model with experiments encourages us to analyse the
molecular binding details from MD simulations. Direct analysis of contacts between ions and
lipids from simulations suggest that the most abundant POPC:\ce{Ca^2+} complex 
has stoichiometry of 2~POPC:1~\ce{Ca^{2+}}.
As shown in Fig.~\ref{fig:cacl-bind} this is in agreement
with the ternary complex model suggested based on head group order parameter
experiments~\cite{altenbach84}.
In addition to the ternary complexes, there also is a non-negligable probability
of one \ce{Ca^{2+}} cross-bridging three POPC molecules.
Technical details of the analysis are in the SI.

@@

The electronic continuum correction is applied here on Lipid14 POPC model \cite{dickson14},
but we expect that the correction can be generalized also for other lipids
and force fields.
The parameters can be used with existing standard nucleic acid and protein force fields, e.g. AMBER-FB15~\cite{Wang2017}. 
We suggest using state of the art water models like OPC3\cite{Izadi16} or OPC \cite{Izadi14},
which yield higher accuracy than the traditional TIP3p water model~\cite{jorgensen83}.
